---
title: "Project_1_final"
author: "Yuchen Shi (ys2901)"
date: "February 3, 2016"
output: pdf_document
---
## Install packages

```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(maps)
```

## Data Preparation 

```{r}
setwd("~/Desktop/Applied Data Science/csv_pus")
pusa <- fread("~/Desktop/Applied Data Science/csv_pus/ss13pusa.csv")
pusb <- fread("~/Desktop/Applied Data Science/csv_pus/ss13pusb.csv")
pus <- rbind(pusa, pusb)

# Here we define chinese data

chinese <- pus%>%
filter(RAC2P==43|RAC2P==44|POBP==207|POBP==209|POBP==240)

# Here we define chinese master students data 

masterchinese.pus = chinese%>%
        filter(SCHL==22)

# Here we interpret our varibles 

chinese$ST <- as.factor(chinese$ST)
chinese$MSP<-as.factor(chinese$MSP)
chinese$SCIENGRLP<-as.factor(chinese$SCIENGRLP)
chinese$SEX<-as.factor(chinese$SEX)
chinese$ESR<-as.factor(chinese$ESR)
levels(chinese$MSP)<-c("married&spouse present","married&spouse absent","Widowed","Divorced","Separated","Never married")
levels(chinese$SCIENGRLP)<-c("Sci","Non-sci")
levels(chinese$SEX)<-c("Male","Female")
levels(chinese$ESR)<-c("empd&work","empd not work","unempd","af&work","af with job but not work","not in labor force")

# Code for state name

levels(chinese$ST) <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut",
"Delaware", "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois",
"Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts",
"Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", 
"New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota",
"Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
"Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia",
"Wisconsin", "Wyoming", "Puerto Rico")

# code for industry

chinese$INDP <- ifelse(chinese$INDP>= 170 & chinese$INDP <= 290, 170, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP>= 370 & chinese$INDP <= 490, 370, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 570 & chinese$INDP<= 770, 570, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 1070 & chinese$INDP <= 3990, 1070, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 4070 & chinese$INDP <= 6390, 4070, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 6470 & chinese$INDP <= 6780, 6470, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP>= 6870 & chinese$INDP <= 7190, 6870, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 7270 & chinese$INDP <= 7790, 7270, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 7860 & chinese$INDP<= 7890, 7860, chinese$INDP)
chinese$INDP<- ifelse(chinese$INDP >= 7970 & chinese$INDP <= 8290, 7970, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 8370 & chinese$INDP <= 8470, 8370, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP %in% c(8660, 8680, 8690), 8370, chinese$INDP) 
chinese$INDP <- ifelse(chinese$INDP >= 8770 & chinese$INDP <= 9290, 8370, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP %in% c(8560, 8570, 8580, 8590, 8670), 8560, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 9370 & chinese$INDP <= 9590, 9370, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 9670 & chinese$INDP<= 9870, 9670, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 9920, 9920, chinese$INDP)
chinese$INDP <- factor(chinese$INDP)
levels(chinese$INDP) <- c("Agriculture, Forestry, Fishing, Hunting", "Mining", "Utilities, Construction", 
              "Manufacturing", "Trade, Logistic", "Information, Communications", "Finance",
              "Professional", "Education", "Health", "Other Services",
              "Arts, Entertainment", "Public Administration", "Military", "Unemployed"
            )

# code for decade

chinese$DECADE <- factor(chinese$DECADE)
levels(chinese$DECADE) <- c("~1950's", "1950's", "1960's", "1970's", "1980's", "1990's", "2000's~")
```


## Where chinese immigrants live or study

```{r}
# prepare data 
all_state <- map_data("state")
data <- as.data.frame(prop.table(table(chinese$ST)))
data$state <- c(sort(tolower(c("district of columbia", state.name))),tolower("Puerto Rico"))
all_state$freq <- data$Freq[match(all_state$region, data$state)]*100

# draw map
p_1 <- ggplot(all_state, aes(x=long, y=lat, group=group)) + 
      geom_polygon(aes(fill=freq), colour="gray78") + 
      scale_fill_gradient(name="Proportion", low="white", high="blue")
p_1 <- p_1 + theme(strip.background = element_blank(),
               strip.text.x     = element_blank(),
               axis.text.x      = element_blank(),
               axis.text.y      = element_blank(),
               axis.ticks       = element_blank(),
               axis.line        = element_blank(),
               panel.background = element_blank(),
               panel.border     = element_blank(),
               panel.grid       = element_blank(),
               legend.position  = "right") +
          xlab("") + ylab("") + ggtitle("Avg. Number of Chinese by State")

p_1
```

We can see that California obviously dominates the place where most chinese live, new york state comes next.

## Average wage of chinese immigrants by state

```{r}
#prepare average income data

average_wage <- chinese %>%
        filter(is.na(WAGP) == F) %>% 
        group_by(ST) %>% 
        summarise(wage = mean(WAGP))
# we don't have Puerto Rico here
state_51 = data[-52,]
state_51$wage = average_wage$wage
all_state$wage = state_51$wage[match(all_state$region,state_51$state)]

#draw map 

p_2 <- ggplot(all_state, aes(x=long, y=lat, group=group)) + 
      geom_polygon(aes(fill=wage), colour="gray78") + 
      scale_fill_gradient(name="wage", low="white", high="blue")
p_2 <- p_2 + theme(strip.background = element_blank(),
               strip.text.x     = element_blank(),
               axis.text.x      = element_blank(),
               axis.text.y      = element_blank(),
               axis.ticks       = element_blank(),
               axis.line        = element_blank(),
               panel.background = element_blank(),
               panel.border     = element_blank(),
               panel.grid       = element_blank(),
               legend.position  = "right") +
          xlab("") + ylab("") + ggtitle("Avg. wage of Chinese by State")

p_2
```

## Chinese Immigrants' working hours by state

```{r}
#prep
all_state <- map_data("state")
data <- as.data.frame(prop.table(table(chinese$ST)))
data$state <- c(sort(tolower(c("district of columbia", state.name))),tolower("Puerto Rico"))
all_state$freq <- data$Freq[match(all_state$region, data$state)]*100

#work hour of chinese masters
work_hour <- chinese %>%
        filter(is.na(WKHP) == F) %>%
        group_by(ST) %>%
        summarise(workhour = mean(WKHP))
state_51 = data[-c(52),]
state_51$workhour = work_hour$workhour
all_state$workhour = work_hour$workhour[match(all_state$region,state_51$state)]

p_work <- ggplot(all_state, aes(x=long, y=lat, group=group))+
        geom_polygon(aes(fill=workhour), color = "gray78") +
        scale_fill_gradient(name="working hour", low="white", high = "blue")
p_work <- p_work + theme(strip.background = element_blank(),
                         strip.text.x     = element_blank(),
                         axis.text.x      = element_blank(),
                         axis.text.y      = element_blank(),
                         axis.ticks       = element_blank(),
                         axis.line        = element_blank(),
                         panel.background = element_blank(),
                         panel.border     = element_blank(),
                         panel.grid       = element_blank(),
                         legend.position  = "right") +
        xlab("") + ylab("") + ggtitle("Avg. working time of Chinese by State")

p_work
```

Form the map, after considering state with enough people, we can see people with highest wage comes from new jersey and texas. We may want to check the income structure in taxes and new jersey.

## Average wage for chinese immigrants with a master degree or higher by state

```{r}
# prepare data 
wage_degree <- chinese %>%
        filter(is.na(WAGP) == F, SCHL>=21) %>% 
        group_by(ST) %>% 
        summarise(wage = mean(WAGP))
state_51d = data[-52,]
# We don't have Puerto Rico, Montana and North Dakoda here
state_51d$wage = wage_degree$wage
all_state$wage = state_51d$wage[match(all_state$region,state_51d$state)]

# draw map

p_3 <- ggplot(all_state, aes(x=long, y=lat, group=group)) + 
      geom_polygon(aes(fill=wage), colour="gray78") + 
      scale_fill_gradient(name="wage", low="white", high="blue")
p_3 <- p_3 + theme(strip.background = element_blank(),
               strip.text.x     = element_blank(),
               axis.text.x      = element_blank(),
               axis.text.y      = element_blank(),
               axis.ticks       = element_blank(),
               axis.line        = element_blank(),
               panel.background = element_blank(),
               panel.border     = element_blank(),
               panel.grid       = element_blank(),
               legend.position  = "right") +
          xlab("") + ylab("") + ggtitle("Avg. wage of Chinese with high degree by State")

p_3
```

We can see California gets darker when we look at people with a high degree, new york and new jersey still dominates the map. we may want to say the income structure in new york, new jersey, and california.

## The wage structure of chinese immigrants in Texas

```{r}
# prepare the data 
TX_structure = chinese %>% 
        filter(ST == "Texas",is.na(WAGP) == F) %>%
        group_by(DECADE, INDP) %>%
        summarise(wage = mean(WAGP))
TX_structure$INDP <- factor(TX_structure$INDP, levels = levels(TX_structure$INDP)[length(levels(TX_structure$INDP)):1])
#Plot the Taxes income structure
ggplot(TX_structure, aes(x=DECADE)) + geom_point(aes(y=INDP, size=wage, colour=INDP)) +
  ggtitle("Avg. Income by Industry and Immigrant Period") +
  guides(colour=FALSE) + theme_minimal()
```

From this graph, we can see that Mining and are a job with a comparetively good salary in each generation; Information and Comunication dominates the job market in 1980's. Health is a goo d job for experienced people but not for youth. 

## Check the working time and salary level for males and females 

```{r}
Male<-subset(masterchinese.pus,masterchinese.pus$SEX=="Male")
Female<-subset(masterchinese.pus,masterchinese.pus$SEX=="Female")
meansal.m<-mean(Male$PINCP,na.rm = T)
meanwkh.m<-mean(Male$WKHP,na.rm = T)
meansal.f<-mean(Female$PINCP,na.rm = T)
meanwkh.f<-mean(Female$WKHP,na.rm = T)
plot(meanwkh.m,meansal.m,xlim = c(25,55),col="red",pch=19,xlab = "mean.workinghour",ylab = "mean.salary")
points(meanwkh.f,meansal.f,pch=19,xlim = c(25,55),col="lightblue")
legend("topright",col = c("red","lightblue"),pch=19,legend = c("Male","Female"),bty="n",cex=0.75)
```

we can see that the average working hour of males and females are very close, 
which indicates the difference in salary is not caused by this factor

## Compare the salary level for Sci and Non-Sci industries and the males' and females' percentage

```{r}
#Industry
salar.ind=
  masterchinese.pus%>%
  group_by(SCIENGP)%>%
  summarise(
    avgsalary=mean(PINCP,na.rm=T)
  )
salar.ind$SCIENGP<-as.factor(salar.ind$SCIENGP)
levels(salar.ind$SCIENGP)<-c("Sci","Non-sci")
plot(salar.ind$avgsalary~salar.ind$SCIENGP)

#Industrys-gender
slices.m<-c(sum(Male$SCIENGP==1),sum(Male$SCIENGP==2))
labels.m<-c("Scientific","Non-scientific")
labels.m <- paste(labels.m, " ", round(slices.m/sum(slices.m)*100,2), "%", sep="")
pie(slices.m,labels.m,main="Industry--Male",col =c("lightgreen","lightyellow"))

slices.f<-c(sum(Female$SCIENGP==1),sum(Female$SCIENGP==2))
labels.f<-c("Scientific","Non-scientific")
labels.f <- paste(labels.f, " ", round(slices.f/sum(slices.f)*100,2), "%", sep="")
pie(slices.f,labels.f,main="Industry--Female",col =c("lightgreen","lightyellow"))
```

we can see that scientific industry earns more than those non-scientific
and also data shows that males are more likelily to be in a scientific industry than females
so the industry difference might be accounted for the salary difference

## Is there gender discrimination ?

```{r}
saldif.sci=
  masterchinese.pus%>%
  filter(SCIENGP=="1")%>%
  group_by(SEX)%>%
  summarise(
    avgsalary=mean(PINCP,na.rm=T)
  )
saldif.sci$SEX<-as.factor(saldif.sci$SEX)
levels(saldif.sci$SEX)<-c("Male","Female")
saldif.nonsci=
  masterchinese.pus%>%
  filter(SCIENGP=="2")%>%
  group_by(SEX)%>%
  summarise(
    avgsalary=mean(PINCP,na.rm=T)
  )
saldif.nonsci$SEX<-as.factor(saldif.nonsci$SEX)
levels(saldif.nonsci$SEX)<-c("Male","Female")

par(mfrow=c(1,2))
plot(saldif.sci,ylim=c(50000,100000),main="sci",cex.axis=0.8,cex.lab=0.8)
plot(saldif.nonsci,ylim=c(50000,100000),main="non-sci",cex.axis=0.8,cex.lab=0.8)
```

As we can see, there do exist gender discrimination at some extent. 
For that under the same education level, close working hour and the same industry, 
there's still much difference in salary between males and females.

## Occupation

```{r}
Occupation=
  Female%>%
  filter(OCCP>0)%>%
  group_by(OCCP)%>%
  summarise(n=n())
  mutate(Occupation,Count=n)%>%
  arrange(desc(Count))
  
occp.sort<- Occupation[order(Occupation$n,decreasing = T),]
occp.sort<- as.data.frame(occp.sort)
occp.sort<- subset(occp.sort,occp.sort$n >30)
barplot(occp.sort$n,names.arg = occp.sort$OCCP,col=terrain.colors(12),cex.names = 0.4)
legend("topright",legend = c("800:ACCOUNTANTS AND AUDITORS","2200:POSTSECONDARY TEACHERS","1020:SOFTWARE DEVELOPERS","2310:ELEMENTARY&MIDDLE TEACHERS","430:MISCELLANEOUS MANAGERS","120:FINANCIAL MANAGERS","1010:COMPUTER PROGRAMMERS","710:MANAGEMENT ANALYSTS","1006:COMPUTER SYSTEMS","5700:SECRETARIES$ASSISTANTS","3255:REGISTERED NURSES","1760:PHYSICAL SCIENTISTS, ALL OTHER"),bty="n",cex=0.7)
```

we're interested in the occupations that are most welcomed by chinese females
as we can see, the occupation which attracts most females is accounting and finance related jobs, which comforms to our assumption.

## Chinese Imigrants' gender distribution on different ages

```{r}
ggplot(chinese, aes(AGEP, group=SEX)) + 
  geom_bar(aes(colour=SEX, fill=SEX), binwidth=1, alpha=0.9) +
  xlab("Age") + ylab("Count") + ggtitle("Chinese Population Age by SEX")
```

## Chinese immigrants' Marital Status distribution 

```{r}
ifmarried <- rep(0,dim(chinese)[1])
for (i in 1:dim(chinese)[1]){
  if (chinese$MAR[i]==1){
    ifmarried[i]= 1
  }
}
ggplot(chinese, aes(AGEP, group=ifmarried)) + 
  geom_bar(aes(colour=ifmarried, fill=ifmarried), binwidth=1, alpha=0.9) +
  xlab("Age") + ylab("Count") + ggtitle("Chinese Population Age by Marital Status")
```

## Chinese immigrants job distribution 

```{r}
counts <- table(chinese$INDP)
barplot(counts,beside=FALSE,main="Industry Distribution", horiz=FALSE,
        col=rainbow(15),xlab="Count",ylab="Industry",names.arg = FALSE,
        legend = c("Agri", "Mining", "Utilities", 
                   "Manufa", "Trade, Logistic", "Information", "Finance",
                   "Professional", "Education", "Health", "Other",
                   "ArtsEnt", "Adm","Military", "Unemployed"))
```






