---
title: "Project_1_final"
author: "Team 1"
date: "February 3, 2016"
output: pdf_document
---

## Introduction

In our first project, we are delighted to present our research on the demographics of Chinese living at the USA. As most of our audience are on the path to be a Chinese Master-degree holder, we explored further about the living conditions and salary level of Chinese master-degree holders.

Through our research, we hope to exploit the income level, standard of living, residence distribution, working condition, gender disparity and marriage. To set off the journey, we began with


## 1: Intall necessary packages for the project

```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(maps)
library(gridExtra)
library(RColorBrewer)
```

## 2: Data Preparation 

As the concentration of the project is only a subset of the total population, filters are mandatory to our research.

```{r}
setwd("~/Desktop/Applied Data Science/csv_pus")
pusa <- fread("~/Desktop/Applied Data Science/csv_pus/ss13pusa.csv")
pusb <- fread("~/Desktop/Applied Data Science/csv_pus/ss13pusb.csv")
pus <- rbind(pusa, pusb)

# Here we define chinese data

chinese <- pus%>%
filter(RAC2P==43|RAC2P==44|POBP==207|POBP==209|POBP==240)

# Here we interpret our varibles 

chinese$ST <- as.factor(chinese$ST)
chinese$MSP<-as.factor(chinese$MSP)
chinese$SCIENGRLP<-as.factor(chinese$SCIENGRLP)
chinese$SEX<-as.factor(chinese$SEX)
chinese$ESR<-as.factor(chinese$ESR)
levels(chinese$MSP)<-c("married&spouse present","married&spouse absent","Widowed","Divorced","Separated","Never married")
levels(chinese$SCIENGRLP)<-c("Sci","Non-sci")
levels(chinese$SEX)<-c("Male","Female")
levels(chinese$ESR)<-c("empd&work","empd not work","unempd","af&work","af with job but not work","not in labor force")
chinese$MSPG <- ifelse(chinese$MSP == 1|chinese$MSP == 2, 1, 0)
chinese$MSPG <-factor(chinese$MSPG)
levels(chinese$MSPG)<-c("Now Married","Other Conditions")
chinese$MSP <- factor(chinese$MSP)
levels(chinese$MSP) <- c("Now married, spouse present", "Now married, spouse absent","Widowed", "Divorced", "Separated", "Never married")
#Add Indicator ESRG 
chinese$ESR <- factor(chinese$ESR)
levels(chinese$ESR) <- c("Employed", "Employed, not at work", "Unemployed", "Employed", "Employed, not at work", "Not in labor force")
chinese$ESRG <- ifelse(chinese$ESR == "Employed", 1, 0)

# Code for state name

levels(chinese$ST) <- c("Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut",
"Delaware", "District of Columbia", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois",
"Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts",
"Michigan", "Minnesota", "Mississippi", "Missouri", "Montana", "Nebraska", "Nevada", 
"New Hampshire", "New Jersey", "New Mexico", "New York", "North Carolina", "North Dakota",
"Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
"Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington", "West Virginia",
"Wisconsin", "Wyoming", "Puerto Rico")

# code for industry

chinese$INDP <- ifelse(chinese$INDP>= 170 & chinese$INDP <= 290, 170, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP>= 370 & chinese$INDP <= 490, 370, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 570 & chinese$INDP<= 770, 570, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 1070 & chinese$INDP <= 3990, 1070, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 4070 & chinese$INDP <= 6390, 4070, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 6470 & chinese$INDP <= 6780, 6470, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP>= 6870 & chinese$INDP <= 7190, 6870, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 7270 & chinese$INDP <= 7790, 7270, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 7860 & chinese$INDP<= 7890, 7860, chinese$INDP)
chinese$INDP<- ifelse(chinese$INDP >= 7970 & chinese$INDP <= 8290, 7970, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 8370 & chinese$INDP <= 8470, 8370, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP %in% c(8660, 8680, 8690), 8370, chinese$INDP) 
chinese$INDP <- ifelse(chinese$INDP >= 8770 & chinese$INDP <= 9290, 8370, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP %in% c(8560, 8570, 8580, 8590, 8670), 8560, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 9370 & chinese$INDP <= 9590, 9370, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 9670 & chinese$INDP<= 9870, 9670, chinese$INDP)
chinese$INDP <- ifelse(chinese$INDP >= 9920, 9920, chinese$INDP)
chinese$INDP <- factor(chinese$INDP)
levels(chinese$INDP) <- c("Agriculture, Forestry, Fishing, Hunting", "Mining", "Utilities, Construction", 
              "Manufacturing", "Trade, Logistic", "Information, Communications", "Finance",
              "Professional", "Education", "Health", "Other Services",
              "Arts, Entertainment", "Public Administration", "Military", "Unemployed"
            )

# code for decade

chinese$DECADE <- factor(chinese$DECADE)
levels(chinese$DECADE) <- c("~1950's", "1950's", "1960's", "1970's", "1980's", "1990's", "2000's~")
chinese$OCCP <- ifelse(chinese$OCCP >= 10 & chinese$OCCP <= 430, 10, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 500 & chinese$OCCP <= 740, 500, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 800 & chinese$OCCP <= 950, 800, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 1005 & chinese$OCCP <= 1240, 1005, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 1300 & chinese$OCCP <= 1560, 1300, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 1600 & chinese$OCCP <= 1965, 1600, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 2000 & chinese$OCCP <= 2060, 2000, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 2100 & chinese$OCCP <= 2160, 2100, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 2200 & chinese$OCCP <= 2550, 2200, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 2600 & chinese$OCCP <= 2920, 2600, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 3000 & chinese$OCCP <= 3540, 3000, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 3600 & chinese$OCCP <= 3655, 3600, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 3700 & chinese$OCCP <= 3955, 3700, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 4000 & chinese$OCCP <= 4150, 4000, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 4210 & chinese$OCCP <= 4250, 4210, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 4300 & chinese$OCCP <= 4650, 4300, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 4700 & chinese$OCCP <= 4965, 4700, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 5000 & chinese$OCCP <= 5940, 5000, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 6005 & chinese$OCCP <= 6130, 6005, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 6200 & chinese$OCCP <= 6765, 6200, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 6800 & chinese$OCCP <= 6940, 6800, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 7000 & chinese$OCCP <= 7630, 7000, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 7700 & chinese$OCCP <= 8965, 7700, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 9000 & chinese$OCCP <= 9750, 9000, chinese$OCCP)
chinese$OCCP <- ifelse(chinese$OCCP >= 9800 & chinese$OCCP <= 9830, 9800, chinese$OCCP)

tempchinese<-chinese%>% 
    filter(OCCP %in% c(9920,5000,3700,800,1600,1005,10,500))
tempchinese$OCCP <- factor(tempchinese$OCCP)
levels(tempchinese$OCCP)<-c("MGR","BUS","FIN","CMM","SCI","PRT","OFF","Unemplyed(broad)")
chinese_Ed<-tempchinese[SCHL>=21]
```

## 3: Age distribution of Chinese

How old are most of the Chinese living in the United States now ?

```{r}
ggplot(chinese, aes(AGEP)) + 
  geom_bar( binwidth=1, alpha=0.9) +
  xlab("Age") + ylab("Count") + ggtitle("Chinese Population Age")
```

Two peaks: The first peak is because of the culture revolution happened in China, US government gives a lot of Chinese immigrants identities. The secodn peak is because of growth of Chinese Economy, more and more people have the chance to study and move to US.

## 4: Are they married ?
How does the pattern of marital status change with time ?

```{r}
ifmarried <- rep(0,dim(chinese)[1])
for (i in 1:dim(chinese)[1]){
  if (chinese$MAR[i]==1){
    ifmarried[i]= 1
  }
}
ggplot(chinese, aes(AGEP, group=ifmarried)) + 
  geom_bar(aes(colour=ifmarried, fill=ifmarried), binwidth=1, alpha=0.9) +
  xlab("Age") + ylab("Count") + ggtitle("Chinese Population Age by Marital Status")
```

Before 25, almost all of Chinese immigrants are not married. Most of the Chinese immigrants choose to marry between the age of 25-30.

## 5: Where do they live ?


```{r}
# prepare data 
all_state <- map_data("state")
data <- as.data.frame(prop.table(table(chinese$ST)))
data$state <- c(sort(tolower(c("district of columbia", state.name))),tolower("Puerto Rico"))
all_state$freq <- data$Freq[match(all_state$region, data$state)]*100

# draw map
p_1 <- ggplot(all_state, aes(x=long, y=lat, group=group)) + 
      geom_polygon(aes(fill=freq), colour="gray78") + 
      scale_fill_gradient(name="Proportion", low="white", high="blue")
p_1 <- p_1 + theme(strip.background = element_blank(),
               strip.text.x     = element_blank(),
               axis.text.x      = element_blank(),
               axis.text.y      = element_blank(),
               axis.ticks       = element_blank(),
               axis.line        = element_blank(),
               panel.background = element_blank(),
               panel.border     = element_blank(),
               panel.grid       = element_blank(),
               legend.position  = "right") +
          xlab("") + ylab("") + ggtitle("Avg. Number of Chinese by State")

p_1
```

California, New York and Texas seem to have the most Chinese. Especially for Colifornia, it seems most Chinese immigrants live there.



## 6: How much do they earn ?

The average standard of living in the US is high, but what about the Chinese? Do they live well and earn a decent living ?

```{r}
#prepare average income data

average_wage <- chinese %>%
        filter(is.na(WAGP) == F) %>% 
        group_by(ST) %>% 
        summarise(wage = mean(WAGP))
# we don't have Puerto Rico here
state_51 = data[-52,]
state_51$wage = average_wage$wage
all_state$wage = state_51$wage[match(all_state$region,state_51$state)]

#draw map 

p_2 <- ggplot(all_state, aes(x=long, y=lat, group=group)) + 
      geom_polygon(aes(fill=wage), colour="gray78") + 
      scale_fill_gradient(name="wage", low="white", high="blue")
p_2 <- p_2 + theme(strip.background = element_blank(),
               strip.text.x     = element_blank(),
               axis.text.x      = element_blank(),
               axis.text.y      = element_blank(),
               axis.ticks       = element_blank(),
               axis.line        = element_blank(),
               panel.background = element_blank(),
               panel.border     = element_blank(),
               panel.grid       = element_blank(),
               legend.position  = "right") +
          xlab("") + ylab("") + ggtitle("Avg. wage of Chinese by State")

p_2
```

The average income varies by state, we can see that New Jersey has the highest average wage. It probabily becuase most of them work in NYC and live in NJ. Texas has comparatively high average wage and it is interesting to look into Taxes.


## 7: How long do they work ?

Chinese are proud to be hard-working, are there some Chinese more hard-working than others?

```{r}
#prep
all_state <- map_data("state")
data <- as.data.frame(prop.table(table(chinese$ST)))
data$state <- c(sort(tolower(c("district of columbia", state.name))),tolower("Puerto Rico"))
all_state$freq <- data$Freq[match(all_state$region, data$state)]*100

#work hour of chinese masters
work_hour <- chinese %>%
        filter(is.na(WKHP) == F) %>%
        group_by(ST) %>%
        summarise(workhour = mean(WKHP))
state_51 = data[-c(52),]
state_51$workhour = work_hour$workhour
all_state$workhour = work_hour$workhour[match(all_state$region,state_51$state)]

p_work <- ggplot(all_state, aes(x=long, y=lat, group=group))+
        geom_polygon(aes(fill=workhour), color = "gray78") +
        scale_fill_gradient(name="working hour", low="white", high = "blue")
p_work <- p_work + theme(strip.background = element_blank(),
                         strip.text.x     = element_blank(),
                         axis.text.x      = element_blank(),
                         axis.text.y      = element_blank(),
                         axis.ticks       = element_blank(),
                         axis.line        = element_blank(),
                         panel.background = element_blank(),
                         panel.border     = element_blank(),
                         panel.grid       = element_blank(),
                         legend.position  = "right") +
        xlab("") + ylab("") + ggtitle("Avg. working time of Chinese by State")

p_work
```

South Dakota, Idaho and Montana residents has proven themselves to be the most hard-working people. Most of them work in the agriculture and forestry industry. They tend to work more and don't earn much.

## 8: Average wage for chinese immigrants with a master degree or higher by state

As most of the our audiences are Master-degree Chinese students, it is important to delight them with our findings in their possible income level after graduation.

```{r}
# prepare data 
wage_degree <- chinese %>%
        filter(is.na(WAGP) == F, SCHL>=21) %>% 
        group_by(ST) %>% 
        summarise(wage = mean(WAGP))
state_51d = data[-52,]
# We don't have Puerto Rico, Montana and North Dakoda here
state_51d$wage = wage_degree$wage
all_state$wage = state_51d$wage[match(all_state$region,state_51d$state)]

# draw map

p_3 <- ggplot(all_state, aes(x=long, y=lat, group=group)) + 
      geom_polygon(aes(fill=wage), colour="gray78") + 
      scale_fill_gradient(name="wage", low="white", high="blue")
p_3 <- p_3 + theme(strip.background = element_blank(),
               strip.text.x     = element_blank(),
               axis.text.x      = element_blank(),
               axis.text.y      = element_blank(),
               axis.ticks       = element_blank(),
               axis.line        = element_blank(),
               panel.background = element_blank(),
               panel.border     = element_blank(),
               panel.grid       = element_blank(),
               legend.position  = "right") +
          xlab("") + ylab("") + ggtitle("Avg. wage of Chinese with high degree by State")

p_3
```

We can see that although, the average wage of Texas gets a little lighter on the map but they are still good as that of California and New York.

## 9: The wage structure of chinese immigrants in Texas

```{r}
# prepare the data 
TX_structure = chinese %>% 
        filter(ST == "Texas",is.na(WAGP) == F) %>%
        group_by(DECADE, INDP) %>%
        summarise(wage = mean(WAGP))
TX_structure$INDP <- factor(TX_structure$INDP, levels = levels(TX_structure$INDP)[length(levels(TX_structure$INDP)):1])
#Plot the Taxes income structure
ggplot(TX_structure, aes(x=DECADE)) + geom_point(aes(y=INDP, size=wage, colour=INDP)) +
  ggtitle("Avg. Income by Industry and Immigrant Period") +
  guides(colour=FALSE) + theme_minimal()
```

From this graph, we can see that Mining and are a job with a comparetively good salary in each generation; Information and Comunication dominates the job market in 1980's. Health industry would be a place to start from ground up and build your exprience in.

## 10: Well-educated Chinese Imigrants' gender distribution on different ages

```{r}
ggplot(chinese_Ed, aes(AGEP, group=SEX)) + 
  geom_bar(aes(colour=SEX, fill=SEX), binwidth=1, alpha=0.9) +
  xlab("Age") + ylab("Count") + ggtitle("Well-educated Chinese Population Age by SEX")
```

It seems that the number of well-educated chinese female immigrants are similar to male's. But for older immigrants, there are more males.

## 11: Gender distribution in each occupation 

We can see each occupation does have different gender distribution.

```{r}
ggplot(chinese_Ed, aes(x=OCCP)) + 
    geom_bar(aes(fill=SEX), position="fill") +ylab("freq")+
    ggtitle("Sex distribution vs Occupation of Chinese with higher education level")
```

For science and computer science, there are a lot more males. For finance and office work, there are a lot more females.

## Is there gender discrimination ?

```{r}
saldif.sci=
  chinese_Ed%>%
  filter(SCIENGP=="1")%>%
  group_by(SEX)%>%
  summarise(
    avgsalary=mean(PINCP,na.rm=T)
  )
saldif.sci$SEX<-as.factor(saldif.sci$SEX)
levels(saldif.sci$SEX)<-c("Male","Female")
saldif.nonsci=
  chinese_Ed%>%
  filter(SCIENGP=="2")%>%
  group_by(SEX)%>%
  summarise(
    avgsalary=mean(PINCP,na.rm=T)
  )
saldif.nonsci$SEX<-as.factor(saldif.nonsci$SEX)
levels(saldif.nonsci$SEX)<-c("Male","Female")

gender = c("male","male","female","female")
salary = c(100125.58,76621.28,76621.28,64367.85)
industry = c("Science","Non-Science","Science","Non-Science")
df = cbind(gender,industry,salary)
df = data.frame(df)
df$salary = salary

ggplot (data = df,aes(industry,salary,fill = gender)) + 
                geom_bar(stat = "identity",position="dodge")+
                xlab("Gender") + ylab("Salary") + 
                ggtitle("Salary for Science Related Work by Gender")
```

As we can see, there is a sign of gender discrimination. 
For people under the same education level, working hour and industry, 
there's still much difference in salary between males and females.












